---
title: "lab4"
author: "Britt Bovbjerg"
date: "3/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data Cleaning: 
```{r}
library(readxl)
library(dplyr)
library(tidyverse)

defense <- read_excel("cyclonesFootball2019.xlsx", sheet="Defensive")
str(defense)

offensive <-read_excel("cyclonesFootball2019.xlsx", sheet="Offensive")

biography <-read_excel("cyclonesFootball2019.xlsx", sheet="Biography")

cols.num <- c("Tackles_Solo","Tackles_ASST", "Tackles_TFL","Tackles_Sack", "Turnover_FF","Turnover_FR","Turnover_INT", "Pass_QBH", "Pass_PB")

defense[cols.num] <- sapply(defense[cols.num],as.numeric)

defClean <- defense

defClean$Name <- as.factor(defClean$Name)

defClean$Opponent_Opponent <-as.factor(defClean$Opponent_Opponent)

str(defClean)
```

```{r}
            
cols.num1 <- c("Rushing_ATT","Rushing_YDS", "Rushing_TD","Receiving_REC", "Receiving_YDS","Receiving_TD","Passing_CMP-ATT", "Passing_YDS", "Passing_TD", "Passing_INT")

offensive[cols.num1] <- sapply(offensive[cols.num1],as.numeric)

offClean <- offensive

offClean$Name <- as.factor(offClean$Name)

offClean$Opponent_Opponent <- as.factor(offClean$Opponent_Opponent)

str(offClean)
```

```{r}

bioClean <- biography

bioClean$Name <- as.factor(bioClean$Name)

bioClean$Weight <- as.numeric(bioClean$Weight)

bioClean <- bioClean %>% 
  separate (col = Height, into= c("Feet", "Inches"),sep='-') %>%
  mutate_at(.vars = vars(Feet, Inches),
            .fun = funs(as.numeric))
 
 bioClean$Heigth <- bioClean$Feet*12 + bioClean$Inches
 
 str(bioClean)

```

Part 2
1)
```{r}
defCleanLong <- defClean %>% pivot_longer(Tackles_Solo:Pass_PB,
    names_to= 'Statistic', values_to="value")

defCleanLong
```
2)
```{r}
ggplot(defCleanLong, aes(x = value)) + geom_histogram(binwidth = .25)  + facet_wrap(~Statistic)
```

It looks like tackles_Sacks, Turnover_FF, Turnover_FR and Turnover_INT are pretty rare.

3)
```{r}

defCleanLong <- defCleanLong %>% 
  filter(Statistic == 'Tackles_Solo') %>%   
  filter(Opponent_Opponent == 'West Virginia' | Opponent_Opponent == 'Kansas State') %>% pivot_wider(names_from='Opponent_Opponent', values_from='value') %>% 
  rename(West_Virginia = 'West Virginia', Kansas_State = 'Kansas State')

defCleanLong %>% ggplot(aes(x=Kansas_State, y=West_Virginia)) + geom_point()

```
The graph doesn't show any signiciant pattern, meaning the team did about as well against Kansas_State as they did against West_Virgina 

Part 2
1)
```{r}
bioClean2 <- bioClean  %>% 
  separate(col = Hometown, into = c('City','State'), sep = ',')

head(bioClean2)
```

2)
```{r}
bioClean2 %>% ggplot(aes(State)) + geom_bar()

table(bioClean2$State)

```

Part 3

1)
```{r}
offBioClean <- left_join(bioClean2, offClean, id = 'Name')

offBioClean %>% ggplot(aes(State)) + geom_bar()

table(offBioClean$State)
```

2)
```{r}
offBioClean %>% 
  filter(!is.na(Opponent_Opponent)) %>%
  ggplot(aes(x= Weight, y = Receiving_YDS)) + geom_point(na.rm = TRUE) + facet_wrap(~Opponent_Opponent)
```
It really doesn't look like there's much of a relationship between weight and recieving yards. There's way too many other factors you would need to look at to make this data mean anything.

3)

```{r}
offBioCleanLong <- offBioClean %>%
  filter(!is.na(Opponent_Opponent)) %>%
  pivot_longer(Rushing_YDS:Rushing_TD,
    names_to= 'Statistic', values_to="value") 
  
  offBioCleanLong %>%
      filter(Name != 'Purdy, Brock')  %>%
      ggplot( aes(x=Opponent_Opponent, y=value))+geom_boxplot()+coord_flip() + facet_grid(~Statistic)
```

```{r}
offBioCleanLong %>%
filter(Name == 'Purdy, Brock') %>% 
ggplot(aes(x=Opponent_Opponent, y=value))+geom_point()+coord_flip() + facet_grid(~Statistic)
```
It looks like Brock Purdy is bettter than the other players at Rushing Yards, but about the same for Rushing_TD. 

4) We wanted to see if weight varied by the state that the player comes from. 
```{r}
ggplot(bioClean2, aes(x=State, y=Weight)) + geom_boxplot()

```
It seems like it doesbecause the means vary quite a bit, with the highest being Illinois and the lowest being Minnasota. 
